---
title: "Lab 01: Data visualization"
subtitle: "Due: Friday, Oct 6 at 11:59pm"
author: ""
date: ""
editor_options: 
  chunk_output_type: console
output:
  pdf_document: default
  html_document: default
---

# Packages

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
``` 

# Exercise 1

```{r ex-1}
data <- mpg
glimpse(data)
```

# Exercise 2

```{r ex-2}
ggplot(data=mpg, mapping=aes(x=cty, y=hwy, color=class))+geom_point()+labs(title="City Miles Per Hour VS Highway Miles Per Gallon (hwy)")
```

Describe the advantages and disadvantages of the plots here (you can delete
these instructions when you're done).
•	geom_point:
Advantages:
Simple and clear way to visualize points on a plot
Easy to control point aesthetics like color, size, shape
Disadvantages:
Points can overlap if there are many observations in the same location
Overlapping points make it hard to see density or count of observations
•	geom_jitter:
Advantages:
Prevents overlapping of points by adding random noise/jitter to the points
Better for dense datasets as it shows underlying patterns and density clearly
Can still see individual observations unlike geom_density()
Disadvantages:
Adds random noise so exact positions are approximations rather than true values
Less precise than geom_point if seeing exact positions is important
More cluttered looking than geom_point if there are many points


# Exercise 3

```{r ex-3}
ggplot(data=mpg, mapping=aes(x=cty, y=hwy, color=class))+geom_jitter()+labs(title="City Miles Per Hour VS Highway Miles Per Gallon (hwy)")
```

# Exercise 4

```{r ex-4}
ggplot(data=mpg, mapping=aes(x=cty, y=hwy, color=class))+geom_jitter()+labs(title="City Miles Per Hour VS Highway Miles Per Gallon (hwy)")+facet_wrap(drv~.)
```

Comment here on what you notice about the relationship between class and city
miles per gallon.
A car's class clearly impacts its city MPG (cty). Larger vehicles like SUVs, trucks and vans get fewer cty MPG than smaller cars. The connection between class and cty MPG seems non-linear, with SUVs lowest and tiny two-seaters highest.

# Exercise 5

```{r ex-5}
ggplot(data=mpg, mapping=aes(x=cty, y=hwy, color=class))+geom_jitter()+labs(title="City Miles Per Hour VS Highway Miles Per Gallon, MN", x= "City Miles Per Gallon (cty)", y="Highway Miles Per Gallon(hwy)")+facet_wrap(drv~.)
```

What do you notice from the bar chart above?
The relationship between city and highway MPG for different vehicle classes. Larger vehicles have lower city MPG than smaller ones. There is a non-linear link between class and city MPG. City MPG also varies significantly within each class.
# Exercise 6

Add a named code chunk here to recreate the plot. Remember, you can insert
code chunks using the "insert chunk" button (a green C with a +) and select `R`. Alternatively, use CMD + OPTION + I (Mac) or CTRL + ALT + I (Windows).
```{r ex-6}
ggplot(data=mpg, aes(x=cty, y=class, color=class))+geom_boxplot()
```

```{r ex-7}
ggplot(data=mpg)+geom_bar(mapping=aes(x=class, fill=drv))+coord_flip()+labs(title="Class VS Drive Train", x="Class")
```

```{r ex-8}
mpg <- mpg %>% 
  rename(vehicle_class = class)
ggplot(data=mpg, mapping=aes(x=displ, y=hwy))+geom_point(size=0.5)+labs(title="Highway mpg VS Displacement",x="Displacement (Litres)", y="Highway mpg")+theme_bw()facet_wrap(~vehicle_class, nrow=2)
```