---
title: "Quiz 01: Data wrangling"
Date: "Wenesday, March 15"
author: ""
date: ""
editor_options: 
  chunk_output_type: console
output:
  pdf_document: default
  html_document: default
---

# Packages

```{r load-packages, message = FALSE}

library(tidyverse)
```

# Data

```{r load-data, message = FALSE}
lego <- read_csv("./lego.csv")
```

# Exercise 1

```{r}
lego_filtered <- lego %>%
filter(!is.na(pieces)) %>%
filter(pieces != 0) %>%
filter(!is.na(retail_price)) %>%
filter(retail_price != 0) %>%
filter(!is.na(year))

```

# Exercise 2

```{r}
lego_filtered %>%
arrange(desc(retail_price)) %>%
slice(1:3)%>%
print(width = Inf)
```

Describe the three most expensive sets here. 
the most expensive 3 lego sets are Millennium Falcon , Connections Kit and Death Star.


# Exercise 3

```{r}
mutate( lego_filtered, "price_per_piece")
lego <- lego %>%
  mutate(price_per_piece = retail_price / pieces)
print(lego, width = Inf)
```

# Exercise 4

```{r}
# Filter out sets with low piece counts
lego_filtered <- lego %>% filter(pieces > 1)

# Create the price_per_piece variable
lego_filtered <- lego_filtered %>%
  mutate(price_per_piece = retail_price / pieces)

# Arrange the dataset in descending order of price_per_piece and select specific columns
lego_sorted <- lego_filtered %>%
  arrange(desc(price_per_piece)) %>%
  select(name, themegroup, theme, pieces, price_per_piece)

# Print the first five rows
head(lego_sorted, 5)
```

Describe what you notice about the sets with the highest price per piece. 

# Exercise 5

```{r}
library(dplyr)

# Filter dataset for "The Lord of the Rings" theme and non-missing retail prices
lego_filtered_lotr <- lego[lego$theme == "The Lord of the Rings" & !is.na(lego$retail_price), ]

# Group the filtered dataset by subtheme
grouped_data <- lego_filtered_lotr %>%
  group_by(subtheme)

# Calculate the cheapest and most expensive sets in each subtheme
result <- grouped_data %>%
  summarize(
    cheapest_set = min(retail_price),
    most_expensive_set = max(retail_price)
  )
result_tibble <- as_tibble(result)
result_tibble
```

# Exercise 6

```{r}
# Remove rows with missing or zero values in retail_price or pieces columns
lego_clean <- lego %>%
  filter(!is.na(retail_price) & !is.na(pieces) & retail_price != 0 & pieces != 0)

# Recalculate yearly trends with the cleaned dataset
yearly_trends <- lego_clean %>%
  group_by(year) %>%
  summarize(
    num_sets = n(),
    median_price_per_piece = median(retail_price / pieces)
  )

# Convert the result to a tibble
yearly_trends <- as_tibble(yearly_trends)

# View the yearly_trends tibble
yearly_trends
```

# Exercise 7

```{r}
library(ggplot2)

# Filter out rows with missing values in the yearly_trends tibble
yearly_trends_filtered <- yearly_trends %>%
  filter(!is.na(year) & !is.na(median_price_per_piece))

# Plotting the median price per piece over time
ggplot(yearly_trends_filtered, aes(x = year, y = median_price_per_piece)) +
  geom_point(aes(size = num_sets), alpha = 0.7, color = "steelblue") +
  scale_size(range = c(3, 10)) +
  labs(x = "Year", y = "Median Price per Piece", size = "Number of Sets") +
  theme_minimal()
```

Comment on what you observe in the plot above. 
The analysis of the median price per piece over time reveals that from 1960 to 1980, the median price per piece was consistently low, with occasional high outliers. However, from 1980 to 1990, there was a shift towards prices closer to the 0.25 threshold. From 1990 to 2020, the median price per piece remained consistently low, without outliers. These findings suggest that LEGO sets have generally maintained an affordable price per piece over the years, with some fluctuations and a trend towards closer alignment with the 0.25 price point.